### 2016-06-28 coverletter.tex
Coverletter and spellings

\documentclass{letter}
\usepackage{txfonts}

\setlength{\textheight}{10.2in} 
\setlength{\topmargin}{-0.9in}

\signature{Simon Cotter}
\address{School of Mathematics \\ Alan Turing Building \\ Oxford
  Road\\Manchester\\ M13 9PL \\ UK \\\textit{email:} simon.cotter@manchester.ac.uk}

\begin{document}

\date{}
\begin{letter}{ }

\opening{To whom it may concern,}

Please find attached a manuscript that we would like you to consider for publication in the SIAM Journal on  Uncertainty Quantification, entitled ``Parallel Adaptive Importance Sampling". We look forward to hearing from you in due course.

\closing{Yours sincerely,}
\end{letter}

\end{document}