### 2018-03-05 cover_letter.tex
Final changes
Final small change, and added the diff file to the response to referees
document.

\documentclass{letter}
\usepackage{txfonts}

\setlength{\textheight}{10.2in} 
\setlength{\topmargin}{-0.9in}

\signature{Dr Simon Cotter (corresponding author), Dr Colin Cotter and Paul Russell}
\address{School of Mathematics \\ Alan Turing Building \\ Oxford
  Road\\Manchester\\ M13 9PL \\ UK \\\textit{email:} simon.cotter@manchester.ac.uk}

\begin{document}

\date{}
\begin{letter}{ }

\opening{Dear Prof. Higdon,}

Please find attached the latest version of our manuscript, ``Parallel Adaptive Importance Sampling''. We have taken onboard the comments of the referees and the associate editor, and we hope that they will be satisfied with the substantial changes and additions that we have made. We very much look forward to hearing from you in due course regarding this paper.

\closing{Yours sincerely,}
\end{letter}

\end{document}