### 2019-01-15 cover_letter.tex
Cover letter added

\documentclass{letter}
\usepackage{txfonts}

\setlength{\textheight}{10.2in} 
\setlength{\topmargin}{-0.9in}

\signature{Dr Simon Cotter (corresponding author), Dr Colin Cotter and
  Dr Paul Russell}
\address{School of Mathematics \\ Alan Turing Building \\ Oxford
  Road\\Manchester\\ M13 9PL \\ UK \\\textit{email:} simon.cotter@manchester.ac.uk}

\begin{document}

\date{}
\begin{letter}{ }

\opening{Dear Prof. Higdon,}

Please find attached the latest version of our manuscript, now
entitled``Ensemble Transport Adaptive Importance Sampling''. We have
taken onboard the comments of the referees and the associate editor,
and we hope that they will be satisfied with the substantial changes
and additions that we have made. We apologise once more for the delay
in completing these revisions. We very much look forward to hearing from you in due course regarding this paper.

\closing{Yours sincerely,}
\end{letter}

\end{document}